Probabilistic Catalogs for Crowded Stellar Fields

### 2013-03-03 paper/ms.tex
Mention methods again

In this paper we have developed and demonstrated a Bayesian approach to
making catalogs from astronomical images in the case where the image contains
only stars (or other point sources). The key idea is that instead of computing
a single catalog, the method creates a posterior probability distribution on
the space of possible catalogs that represents our state of knowledge about
the presence and properties of objects in the image. When this is done, the
uncertainties in the imaging are accurately propagated through to scientific
conclusions, for example about the luminosity function of the stars.
This approach was contrasted with the results from the standard \sex~software.
For the bright sources the results were essentially consistent, however the
Bayesian approach was more successful at modelling the distribution of faint
stars. Of course, the Bayesian method is much more computationally intensive,
which is a significant issue in practice. To build a more complete picture of
when our approach is necessary in practice, it will need to be tested against
a wider variety of alternative methods.

We note that there are many limitations to the model presented in this paper,
some of which will be important to relax when applying it to real data.
In principle, our model should be a model of the physical state of the universe,
and not a simple model where the only stellar properties are a 2-D position and
a flux. Another limitation is that we have not considered multi-epoch or multi-band
imaging. In the former case, PSF variations and stellar motions may be relevant
\citep{lang}, and in the latter, a model for the spectral energy distributions
of the stars will need to be considered: essentially, the luminosity function
will need to be a probability distribution over more than one dimension.

>>>>

In this paper we have developed and demonstrated a Bayesian approach to
making catalogs from astronomical images in the case where the image contains
only stars (or other point sources). The key idea is that instead of computing
a single catalog, the method creates a posterior probability distribution on
the space of possible catalogs that represents our state of knowledge about
the presence and properties of objects in the image. When this is done, the
uncertainties in the imaging are accurately propagated through to scientific
conclusions, for example about the luminosity function of the stars.
This approach was contrasted with the results from the standard \sex~software.
For the bright sources the results were essentially consistent, however the
Bayesian approach was more successful at modelling the distribution of faint
stars. Of course, the Bayesian method is much more computationally intensive,
which is a significant issue in practice. To build a more complete picture of
when our approach is necessary in practice, it will need to be tested against
a wider variety of alternative methods  such as DAOPHOT \citep{1987PASP...99..191S}
and DOLPHOT \citep{dolphot}.

We note that there are many limitations to the model presented in this paper,
some of which will be important to relax when it is applied to real data.
In principle, our model should be a model of the physical state of the universe,
and not a simple model where the only stellar properties are a 2-D position and
a flux. Another limitation is that we have not considered multi-epoch or multi-band
imaging. In the former case, PSF variations and stellar motions may be relevant
\citep{lang}, and in the latter, a model for the spectral energy distributions
of the stars will need to be considered: essentially, the luminosity function
will need to be a probability distribution over more than one dimension.
